\section{Discussion}

The experimental results indicate that architectural complexity correlates positively with multi-label recommendation quality up to a point, beyond which performance gains plateau. FusionNet's design choice of maintaining separate encoding pathways for symptom text and structured patient features proves particularly well-suited to the outpatient setting, where demographic and comorbidity signals carry meaningful clinical weight alongside free-text complaints. The explicit fusion of these two modalities---rather than simple concatenation at the embedding level, as in the MLP and CNN baselines---enables the model to learn richer cross-modal interactions, a capacity reflected in its superior subset accuracy and Jaccard index.

The consistently high ROC-AUC scores across all architectures suggest that the underlying symptom--medication signal present in the synthetic dataset is learnable even by comparatively shallow models. Nevertheless, the substantial gap in subset accuracy between the MLP (0.4210) and FusionNet (0.6455) underscores that correctly recovering the complete medication set for a given patient---rather than predicting individual drug labels in isolation---is a considerably harder problem, one that benefits from sequential or attention-based inductive biases. This distinction carries direct clinical significance: a system that correctly identifies most medications but omits a single critical drug or includes one unsafe candidate is of limited practical utility.

The uniformly low macro-averaged F1 scores observed across all models reflect a well-documented challenge in medication recommendation: the long-tail distribution of drug classes. Infrequent medications, which may be of particular clinical importance in cases involving rare comorbidities or drug allergies, are systematically under-predicted. This limitation is partly an artefact of the synthetic data generation process, in which drug assignment is governed by disease--drug mappings rather than authentic physician behavior, and may therefore underrepresent the nuance of real prescribing decisions. Addressing this distributional imbalance through class-weighted loss functions, oversampling strategies, or prototype-based learning for rare drug classes constitutes an important direction for future work.

SelfAttention's comparatively strong PR-AUC and Precision@$K$ performance relative to its subset accuracy suggests that attention mechanisms are particularly effective at surfacing the most clinically relevant candidates at the top of the ranked list, even when the full predicted set is not recovered exactly. This property is desirable in a clinical decision support context, where a ranked shortlist presented to the prescribing clinician is often more actionable than a fixed predicted label set. Integrating self-attention as a ranking re-scorer applied over FusionNet's initial predictions represents a promising avenue for architectural refinement.

A notable limitation of the current evaluation is the absence of an explicit DDI rate comparison across models at inference time. The safety filtering layer described in the methodology was applied during dataset construction rather than as a post-hoc inference component in the present experiments. Future work should decouple the safety filter from data generation and measure the raw DDI rate of model outputs both before and after filtering, enabling direct quantification of the safety benefit conferred by the graph-based interaction layer. Furthermore, all experiments were conducted on synthetically generated data, and performance on real-world clinical corpora may differ considerably owing to the distributional gap between template-constructed symptom text and authentic clinical documentation.

\section{Conclusion}

This paper presented a structured survey of twenty deep learning approaches to medication recommendation, identified persistent gaps in outpatient generalization, safety integration, and interpretability, and proposed a hybrid architecture designed to address these limitations. Five neural models of increasing complexity were implemented and evaluated on a synthetic yet clinically grounded multi-label dataset, constructed by integrating symptom--disease mappings, drug--disease associations, DDI constraints, and Synthea-derived demographic records.

FusionNet, which combines a sequential text encoder with a dedicated structured feature branch, achieved the strongest overall performance across Jaccard index, micro-F1, and subset accuracy, demonstrating the value of modality-aware feature fusion for outpatient medication recommendation. SelfAttention exhibited competitive top-$k$ ranking performance, indicating its suitability for clinician-facing shortlist generation. All architectures attained high ROC-AUC scores, confirming that the symptom-to-medication mapping signal is robustly learnable across model families, while uniformly low macro-F1 scores identified rare drug prediction as a persistent open challenge.

The proposed framework incorporates a DDI-aware safety filter and a SHAP-based interpretability layer to address clinical trust and patient safety requirements beyond raw predictive performance. Collectively, these components establish a foundation for a lightweight, safety-aware, and interpretable medication recommendation system suitable for primary-care decision support under the data-sparse conditions characteristic of outpatient practice.

Future work will pursue evaluation on real-world clinical datasets, incorporation of class-imbalance mitigation strategies for rare drug classes, extension of the safety filter to runtime inference with quantifiable DDI rate reduction, and investigation of larger pretrained clinical language models as drop-in symptom encoders within the FusionNet backbone.
