\section{Introduction}
Medication prescribing in clinical practice is an inherently complex decision-making process, requiring the concurrent evaluation of therapeutic efficacy, patient-specific contraindications, dosage considerations, and potential adverse drug interactions. Clinicians must navigate a substantial pharmacological search space while accounting for patient comorbidities, demographic factors, and interaction-related risks. This cognitive complexity contributes to observable variability in prescribing behavior and raises the likelihood of suboptimal or clinically unsafe medication combinations. The widespread adoption of electronic health records (EHRs) has created new opportunities for data-driven approaches to assist in medication recommendation and broader clinical decision support.

Deep learning methods have demonstrated considerable capacity for modeling the complex, high-dimensional structure of healthcare data. Early contributions, such as \textit{Doctor AI} \cite{choi2016doctorai}, employed recurrent neural networks (RNNs) over longitudinal EHR sequences to predict future diagnoses and medication categories. By deriving temporal patient representations from successive clinical visits, these models captured disease progression dynamics and yielded improvements in multi-label prediction. Complementarily, \textit{RETAIN} \cite{choi2016retain} introduced a reverse-time attention mechanism that enhanced model interpretability without sacrificing predictive performance. Through the assignment of attention weights to prior visits and clinical variables, RETAIN offered clinicians greater transparency into model reasoning---an important consideration for the practical deployment of deep learning systems in healthcare contexts.

Subsequent research has extended these foundations by incorporating graph-based representations and external medical knowledge to improve both robustness and safety. \textit{GAMENet} \cite{shang2019gamenet} combined patient history encoding with a Graph Convolutional Network (GCN) applied over drug--drug interaction (DDI) graphs, explicitly modeling adverse interaction structures to reduce unsafe recommendation combinations while preserving accuracy. Related work, including \textit{SafeDrug} \cite{lee2021safedrug}, employed molecular graph encoders to capture chemical substructure information and support DDI-aware prediction. Collectively, these contributions demonstrate how deep architectures integrating temporal modeling, graph neural networks, and domain knowledge can advance both the effectiveness and safety of medication recommendation systems.

Nevertheless, the majority of existing approaches presuppose access to rich longitudinal EHR data, of the kind typically available in tertiary-care or inpatient settings. Outpatient and primary-care environments present a markedly different data landscape, often characterized by sparse records, limited visit histories, and brief free-text symptom descriptions. In many encounters, a single visit with minimal structured history constitutes the entirety of available information, which substantially constrains the utility of sequence-dependent models. Moreover, the outpatient context demands that safety considerations be given particular prominence, ensuring that recommended drug combinations minimize interaction risks even in the absence of comprehensive longitudinal context.

This work addresses these challenges through a focused survey of deep learning-based medication recommendation systems, with emphasis on safety modeling, interpretability, and suitability for data-sparse outpatient settings. Drawing on identified gaps, we propose a hybrid deep learning architecture that integrates attention-based symptom encoding, structured feature fusion, and DDI-aware safety filtering. The overarching objective is to develop a medication recommendation framework that appropriately balances predictive performance, safety constraints, and interpretability for application in primary-care clinical decision support.
