\section{Literature Review}

Recent research on medication recommendation systems has increasingly leveraged deep learning to address the complexity, scale, and safety requirements of clinical decision-making. Early approaches focused on sequential modeling of electronic health records (EHRs) through recurrent neural networks, with the aim of capturing temporal disease progression and prescribing patterns. Subsequent developments have incorporated attention mechanisms, graph neural networks, and knowledge-based constraints to improve interpretability and reduce adverse drug--drug interactions. In parallel, transformer-based language models and transfer learning techniques have demonstrated strong performance in extracting clinically meaningful representations from short, unstructured symptom text. Collectively, these trends reflect a broad shift toward hybrid architectures that combine temporal modeling, structured medical knowledge, and explainable AI to produce safer and more reliable medication recommendations, particularly in data-sparse clinical settings.


\begin{enumerate}[label=\textbf{\arabic*.}]
    \item \textbf{DP-CRNN: CNN--RNN for pre-diagnosis (Zhou et al.)} \\
    Zhou et al. propose a double-pooling convolutional-recurrent architecture that first extracts $n$-gram semantics via convolutional filters, subsequently models sequential patterns through recurrent layers, and applies a secondary pooling step to emphasize salient temporal features. While the architecture is well-suited to consumer-style free-text inputs, it does not extend to multi-drug prescription tasks.

    \item \textbf{Deep Patient (Miotto et al.)} \\
    Miotto et al. train stacked denoising autoencoders on large EHR collections to produce latent patient representations suitable for downstream prediction. The work demonstrates that unsupervised representation learning can recover clinically meaningful structure from raw records; however, temporal ordering is not explicitly modeled.

    \item \textbf{Doctor AI (Choi et al.)} \\
    Doctor AI employs recurrent neural networks over sequences of clinical visits to forecast future diagnoses and medications. The model establishes the value of sequential EHR modeling, though interpretability remains limited and drug--drug interactions are not explicitly addressed.

    \item \textbf{RETAIN (Choi et al.)} \\
    RETAIN introduces a reverse-time attention mechanism that produces interpretable importance scores over historical clinical events. Relative to conventional recurrent architectures, it offers improved transparency; however, the model was not designed to enforce multi-drug safety constraints.

    \item \textbf{GAMENet (Shang et al.)} \\
    GAMENet combines a memory-augmented patient encoder with graph-augmented drug representations to recommend medication sets while reducing harmful drug combinations. Its dependence on rich longitudinal EHR data limits direct applicability to sparse outpatient contexts.

    \item \textbf{SafeDrug (Lee et al.)} \\
    SafeDrug employs molecular graph encoders to capture chemical substructures, integrating these representations with clinical context to reduce adverse interaction risk. While the molecular modeling approach meaningfully constrains unsafe predictions, it introduces considerable computational overhead.

    \item \textbf{GRAM (Choi et al.)} \\
    GRAM augments clinical code representations by incorporating hierarchical medical ontology structure---such as the ICD taxonomy---via graph attention, yielding improved performance on infrequent diagnostic codes.

    \item \textbf{LEAP (Zhang et al.)} \\
    LEAP frames multi-drug prescription as a constrained optimization problem balancing therapeutic efficacy and interaction safety. The approach models drug combinations directly, but relies on long-term outcome labels that are seldom available in outpatient datasets.

    \item \textbf{Molecular and substructure-based models (MoleRec)} \\
    Molecular-focused methods apply message-passing neural networks to learn drug representations from chemical graph structures, supporting more principled interaction prediction. This additional representational fidelity comes at the cost of increased model complexity.

    \item \textbf{ClinicalBERT and transformer-based models} \\
    Clinical adaptations of the BERT architecture yield strong contextual embeddings for clinical notes and demonstrate particular effectiveness in capturing symptom semantics. Their primary limitation is substantial computational cost at inference.

    \item \textbf{Memory-augmented networks} \\
    These architectures maintain external memory banks of prior visit representations to inform current recommendations. Their effectiveness is contingent on sufficient patient history, which is frequently absent in outpatient settings.

    \item \textbf{Hierarchical Attention Networks (HAN)} \\
    HAN processes clinical text at multiple levels of granularity (word $\rightarrow$ sentence) and identifies relevant passages for interpretability, a property of particular value in clinician-facing explanation systems.

    \item \textbf{GRU-D (Che et al.)} \\
    GRU-D incorporates missingness masking and temporal decay mechanisms to handle the irregular sampling characteristic of clinical time series, improving robustness for laboratory- and vital-sign-based prediction tasks.

    \item \textbf{Polypharmacy side-effect modeling (Zitnik et al.)} \\
    This line of work models adverse outcomes of drug combinations through graph neural networks applied to drug--drug interaction networks, enabling direct probabilistic estimation of interaction severity.

    \item \textbf{Transfer learning for clinical NLP} \\
    Pretrained language models demonstrate consistent performance gains under low-resource conditions, supporting more reliable symptom-to-diagnosis mapping from limited labeled datasets.

    \item \textbf{Multi-task and multi-label models} \\
    These frameworks jointly predict diagnoses, clinical urgency, and medication sets by exploiting shared representational structure across tasks. Careful task weighting is necessary to prevent negative transfer between objectives.

    \item \textbf{Reinforcement learning for treatment policies} \\
    Reinforcement learning approaches optimize sequential treatment decisions with respect to long-term outcome criteria. Such methods require longitudinal reward signals and present non-trivial safety risks if deployed without appropriate constraints.

    \item \textbf{Explainable AI (LIME and SHAP)} \\
    Model-agnostic local explanation frameworks generate instance-level rationales for individual predictions. The integration of graph- or rule-based explanations alongside these methods tends to yield stronger clinical plausibility and stakeholder trust.

    \item \textbf{Knowledge graph embedding approaches} \\
    These methods situate diseases, drugs, and clinical procedures within a structured latent space, enabling reasoning over relational medical knowledge. Their practical deployment requires careful ontology curation and cross-source harmonization.

    \item \textbf{Large clinical transformer architectures} \\
    Models trained on heterogeneous clinical signals---combining unstructured notes with structured EHR data---achieve state-of-the-art predictive performance across a range of tasks, though they demand substantial computational infrastructure and raise significant data privacy considerations.
\end{enumerate}

In summary, the surveyed literature reflects considerable progress in applying deep learning to medication recommendation, spanning sequential models, attention-based architectures, graph neural networks, molecular representations, and large-scale transformers. While many approaches attain strong predictive performance, the majority rely on dense longitudinal EHR data and are tailored to inpatient or intensive-care settings. Safety-aware models that incorporate drug--drug interaction knowledge have emerged as an important subfield, yet frequently at the cost of increased architectural complexity or heightened data requirements. Interpretability and outpatient generalization remain substantive open challenges. These limitations underscore the need for lightweight, safety-aware, and explainable deep learning frameworks capable of operating effectively under the sparse data conditions characteristic of primary-care environments.


From the analyses above, several persistent gaps can be identified:
\begin{itemize}
    \item \textbf{Outpatient generalization:} The majority of safety-aware methods presuppose lengthy EHR histories; outpatient inputs are characteristically brief and sparse.
    \item \textbf{Integrated safety and explainability:} Few existing systems simultaneously provide DDI-aware recommendations and human-interpretable rationales within a unified framework.
    \item \textbf{Dosage and demographic adaptation:} Most approaches recommend drug identities without offering patient-specific dose ranges adjusted for weight, age, or renal function.
\end{itemize}
