% -----------------------------------------------
\section{Introduction}
% -----------------------------------------------

\begin{frame}{The Problem with Medication Prescribing}
  \begin{itemize}
    \item Prescribing requires balancing:
    \begin{itemize}
      \item Therapeutic efficacy
      \item Contraindications \& dosage
      \item Drug--drug interactions (DDIs)
      \item Patient comorbidities \& demographics
    \end{itemize}
    \pause
    \item EHRs create new opportunities for \textbf{data-driven clinical decision support}
  \end{itemize}
  \note{Set the stage -- prescribing is hard, mistakes happen, and EHRs give us the data to do something about it.}
\end{frame}

\begin{frame}{Surveyed Studies (1--7)}
\small
\begin{tabularx}{\textwidth}{@{}p{0.36\textwidth} l l X@{}}
\toprule
\textbf{Study / Year} & \textbf{Arch.} & \textbf{Safety} & \textbf{Limitation} \\
\midrule
DP-CRNN (2021) \cite{zhou2021dpcrnn} & CNN+RNN & No & Not drug-centric \\
Deep Patient (2016) \cite{miotto2016deep} & Autoenc. & No & Static reps \\
Doctor AI (2016) \cite{choi2016doctorai} & RNN & No & Low interpretability \\
RETAIN (2016) \cite{choi2016retain} & Attn-RNN & No & Not drug-aware \\
GAMENet (2019) \cite{shang2019gamenet} & GNN+Mem & Yes & Dense EHR needed \\
SafeDrug (2021) \cite{lee2021safedrug} & Mol-GNN & Yes & Compute heavy \\
GRAM (2017) \cite{choi2017gram} & Ontology GAT & Partial & Ontology dependent \\
\bottomrule
\end{tabularx}
\end{frame}

\begin{frame}{Surveyed Studies (8--14)}
\small
\begin{tabularx}{\textwidth}{@{}p{0.36\textwidth} l l X@{}}
\toprule
\textbf{Study / Year} & \textbf{Arch.} & \textbf{Safety} & \textbf{Limitation} \\
\midrule
LEAP (2017) \cite{zhang2017leap} & Opt+Seq & Partial & Needs labels \\
MoleRec (2022–23) \cite{moleRec} & GNN & Yes & Complex training \\
ClinicalBERT (2019) \cite{alsentzer2019clinicalbert} & Transformer & No & Pretrain cost \\
MemoryNet (2019) \cite{wang2019memory} & Memory NN & No & Needs history \\
HAN (2016) \cite{yang2016han} & Hier. Attn & No & Complex \\
GRU-D (2018) \cite{che2018grud} & GRU-Decay & No & Vitals focused \\
Polypharmacy GNN (2018) \cite{zitnik2018polypharmacy} & GNN & Yes & Graph sensitive \\
\bottomrule
\end{tabularx}
\end{frame}

\begin{frame}{Surveyed Studies (15--20)}
\small
\begin{tabularx}{\textwidth}{@{}p{0.36\textwidth} l l X@{}}
\toprule
\textbf{Study / Year} & \textbf{Arch.} & \textbf{Safety} & \textbf{Limitation} \\
\midrule
Transfer Clinical NLP (2020) \cite{lee2020transfer} & Pretrained & No & Domain shift \\
Multi-task CDSS (2019) \cite{harutyunyan2019multitask} & Multi-DNN & Partial & ICU focused \\
RL Treatments (2018) \cite{komorowski2018rl} & RL & Partial & Reward design \\
LIME / SHAP \cite{ribeiro2016lime} & Explainers & No & Approximate \\
KG Embeddings \cite{kg_embeddings} & KG models & Yes & KG curation \\
Clinical Transformers (2022) \cite{transformer_clinical2022} & Transformer & Partial & Resource heavy \\
\bottomrule
\end{tabularx}
\end{frame}

\begin{frame}{The Gap: Outpatient \& Data-Sparse Settings}
  \begin{itemize}
    \item Existing models assume \textbf{rich longitudinal EHR data} (inpatient/tertiary care)
    \pause
    \item Outpatient reality:
    \begin{itemize}
      \item Sparse records, limited visit history
      \item Brief free-text symptom descriptions
      \item Often just \textbf{a single visit}
    \end{itemize}
    \pause
    \item Safety is \emph{even more} critical without longitudinal context
  \end{itemize}
  \note{This is the key motivation slide. Emphasize that sequence-dependent models just don't work here -- and safety can't be an afterthought.}
\end{frame}

\begin{frame}{Our Approach}
  \begin{itemize}
    \item A hybrid deep learning architecture for \textbf{outpatient medication recommendation}
    \pause
    \item Key components:
    \begin{itemize}
      \item Attention-based symptom encoding
      \item Structured feature fusion
      \item DDI-aware safety filtering
    \end{itemize}
    \pause
    \item Goals: \textbf{performance}, \textbf{safety}, and \textbf{interpretability}
  \end{itemize}
  \note{Transition slide into the rest of the presentation. You can forward-reference the architecture diagram here if it comes up in the next section.}
\end{frame}
